---
title: "Comparison of Algorithms Across Business Lines"
output:
  html_document:
    df_print: paged
---


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}
source('report_fun.R')


# This section sets up the choice of business lines to be analyzed as well as the MCMC parameters

iter = 1200
warmup = 500
chains = 8
adapt_delta = 0.9
max_treedepth = 12
business_lines = c("comauto", "medmal", "othliab","ppauto","prodliab", "wkcomp")
#business_lines = c("prodliab")


#limit number of cases per line
if (F){
  table(data[, c("line", "GRCODE")])
  company_max = 2
  for (bus in business_lines){
    all_lines = unique(data$GRCODE[data$line == bus])
    data = data %>%
      filter(line != bus | GRCODE %in% tail(all_lines,company_max))
      #filter(line != bus | GRCODE %in% all_lines[4])
  } 
  (table(data$line))
}
```



# Introduction

This notebook provides the code for the paper "Gaussian Process Models for Incremental Loss Ratios" by Michael Ludkovski and Howard Zail.


## Models Analyzed

The notebook tests key algorithms across the different proerty and casualty business lines for projecting paid loss triangles by "completing the square".  The algorithms tested are:

**LDF Models**

* Mack Chain Ladder
* Bootstrapped Chain Ladder

**Incremental Claim Models**

* GP on Incremental Loss Ratios with no constraints on GP using a compound kernel (all models below use compond kernel)
* GP on Incremental Loss Ratios  with hurdle model
* GP on Incremental Loss Ratios  with hurdle model and virtual data points


## Stan Programs

This notebook runs Stan programs iteratively.  It is important for speed purposes that each of the Stan programs are pre-compiled. Once the code is compiled, Stan stores a compiled version.  But sometimes this routine fails if the compilation is done as part of a subroutine. Thus it is important that the following code be run once MANUALLY before running the rest of the program.  

```{r}
stanMod <- stan_model(file=paste0(stan_dir, "gp_compound_mean_het_05.stan"))
stanMod <- stan_model(file=paste0(stan_dir,  "gp_hurdle_phiapprox_05.stan"))
```

### Tuning of Rho

The priors for rho are determined using the Betancourt method.  This fits an inverse gamma distribution to rho, such that the each of probability that the length scale of rho is less than the minimum or or greater than the maximum is $\alpha \%$.  In the report, we set $\alpha=0.001$ (the tail probability on each side).  We work with the scaled accident year and development lag input dimensions, which have the same minimum and maximum length scales.  

##### Tuning for ILR Models

ILR models train on 10x10 squares of incremental claims

```{r echo=FALSE}

# Options:  
alpha = 0.001

# Options:  Preferable for plain model
#alpha = 0.1



scaled_development_lag = c(scale(1:10))

min_length_scale = abs(scaled_development_lag[1]-scaled_development_lag[2]) # choose any two sequential points
max_length_scale = abs(max(scaled_development_lag)-min(scaled_development_lag)) # choose any two sequential points

prior_rho = rho_prior_tune(tail_size = alpha, lower_ls = min_length_scale, upper_ls = max_length_scale)

prior_rho = prior_rho$par
```

* The minimum length scale is: `r min_length_scale`
* The maximum length scale is: `r max_length_scale`
* The inverse gamma prior parameters for rho (for each of the AccidentYear and DevelopmentLag dimensions are) are: `r round(prior_rho,3)`



# The Algorithms


## Mack Chain Ladder


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

res_mack_ladder = all_business_line_calcs(business_lines = business_lines, data = data, projection_fun = mack_ladder, iter = (iter-warmup)*chains,  
                                     include_rank_tests = TRUE)


save(res_mack_ladder , file = paste0(output_dir, "report_algos_mack_ladder.RData"))
load(file = paste0(output_dir, "report_algos_mack_ladder.RData"))


```

```{r  echo=FALSE}
print_loss_results(res_mack_ladder)
```


## Bootstrapped Chain Ladder

```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

res_boot_ladder = all_business_line_calcs(business_lines = business_lines, 
                                          data = data, projection_fun = boot_ladder, iter = (iter-warmup)*chains,
                                          include_rank_tests = TRUE)


#save(res_boot_ladder , file = paste0(output_dir, "report_algos_boot_ladder.RData"))
#load(file = paste0(output_dir, "report_algos_boot_ladder.RData"))


```

```{r  echo=FALSE}
print_loss_results(res_boot_ladder)
```



## GP on Incremental Loss Ratios with no constraints on GP using a compound kernel


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

res_gp_inc = all_business_line_calcs(business_lines = business_lines, data = data, 
                                     projection_fun = stan_incremental_loss_ratio3,  
                                     calc_ranks = T, include_rank_tests = TRUE, verbose = FALSE,
                                     iter = iter, warmup=warmup, chains=chains, adapt_delta = adapt_delta, 
                                     max_treedepth = max_treedepth,
                                     prior_eta = 1, 
                                     prior_rho = prior_rho,
                                     prior_theta = 1, prior_sigma = 1, take_logs = FALSE, 
                                     mu=0,
                                     formula_x = y_scaled ~ AccidentYear_scaled + DevelopmentLag_scaled -1,
                                     formula_H = y_scaled ~ AccidentYear_scaled + log(DevelopmentLag) -1,
                                     stanFile =  "gp_compound_mean_het_05.stan")

save(res_gp_inc , file = paste0(output_dir, "report_algos_gp_inc.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc.RData"))


```




```{r,echo=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
print_loss_results(res_gp_inc)
```


## GP on Incremental Loss Ratios with Hurdle Model with Vector mu


```{r echo=FALSE, message=FALSE, include=FALSE, results='hide'}

res_gp_inc_hurdle = all_business_line_calcs(business_lines = business_lines, data = data, 
                                            projection_fun = stan_incremental_loss_ratio3,  
                                     calc_ranks = T, verbose = FALSE,
                                     iter = iter, warmup=warmup, chains=chains, adapt_delta = adapt_delta, 
                                     max_treedepth = max_treedepth,
                                     prior_sigma = 1, mu=10:1 / 10 -.5, 

                                     prior_rho = prior_rho,
                                     prior_eta = 1, prior_theta = 1, take_logs = FALSE, 
                                     formula_x = y_scaled ~ AccidentYear_scaled + DevelopmentLag_scaled -1,
                                     formula_H = y_scaled ~ AccidentYear_scaled + log(DevelopmentLag) -1,
                                     stanFile = "gp_hurdle_phiapprox_05.stan")

save(res_gp_inc_hurdle , file = paste0(output_dir, "report_algos_gp_inc_hurdle.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_hurdle.RData"))

```




```{r,echo=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
print_loss_results(res_gp_inc_hurdle)

```



## GP on Incremental Loss Ratios with Hurdle Model with Vector mu, with Virtual Data 


```{r echo=FALSE, message=FALSE, include=FALSE, results='hide'}
res_gp_inc_hurdle_virtual = all_business_line_calcs(business_lines = business_lines, data = data, 
                                                    projection_fun = stan_incremental_loss_ratio3,  
                                     calc_ranks = T, verbose = FALSE,
                                     iter = iter, warmup=warmup, chains=chains, adapt_delta = adapt_delta, 
                                     max_treedepth = max_treedepth,
                                     #prior_rho = c(5.566019, 5.570317), 
                                     #prior_rho = c(9.4,8.4),
                                     prior_rho = prior_rho, 
                                     prior_sigma = 1, 
                                     mu=10:0 / 10 -.5, 
                                     prior_eta = 1, prior_theta = 1, take_logs = FALSE, 
                                     formula_x = y_scaled ~ AccidentYear_scaled + DevelopmentLag_scaled -1,
                                     formula_H = y_scaled ~ AccidentYear_scaled + log(DevelopmentLag) -1,
                                     virtual_data = T, virtual_point = -0.3, virtual_accident_year = 1988:1997,
                                     virtual_development_lag = 11,
                                     stanFile = "gp_hurdle_phiapprox_05.stan")
save(res_gp_inc_hurdle_virtual , file = paste0(output_dir, "report_algos_gp_inc_hurdle_virtual.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_hurdle_virtual.RData"))

```



```{r,echo=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
print_loss_results(res_gp_inc_hurdle_virtual)
```







