---
title: "Multi Company Model"
output: html_notebook
---


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}
source('loss_multi_company.R')

iter = 800
warmup = 500
iter_multi=800
warmup_multi=200

chains = 8

adapt_delta = 0.9
max_treedepth = 11
#business_lines = c("comauto", "medmal", "othliab","ppauto","prodliab", "wkcomp")
business_lines = c("wkcomp")
```


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

# Select a subset of policies

set.seed(666)
company_max = 5
#company_max = Inf
# for (bus in business_lines){
#   companies = unique(data$GRCODE[data$line == bus])
#   company_count = min(length(companies), company_max)
#   all_lines = sample(companies, size = company_count, replace = FALSE) 
#   data = data %>%
#     filter(line != bus | GRCODE %in% sample(all_lines, size = company_max))
# }


for (bus in business_lines){
  all_lines = unique(data$GRCODE[data$line == bus])
  data = data %>%
    filter(line != bus | GRCODE %in% head(all_lines,company_max))
} 


```


### Tuning of Rho

The priors for rho are determined using the Betancourt method.  This fits an inverse gamma distribution to rho, such that the each of probability that the length scale of rho is less than the minimum or or greater than the maximum is $\alpha \%$.  In the report, we set $\alpha=0.001$ (the tail probability on each side).  We work with the scaled accident year and development lag input dimensions, which have the same minimum and maximum length scales.  

##### Tuning for ILR Models

ILR models train on 10x10 squares of incremental claims

```{r echo=FALSE}
alpha = 0.001

scaled_development_lag = c(scale(1:10))

min_length_scale = abs(scaled_development_lag[1]-scaled_development_lag[2]) # choose any two sequential points
max_length_scale = abs(max(scaled_development_lag)-min(scaled_development_lag)) # choose furthest  points

prior_rho = rho_prior_tune(tail_size = alpha, lower_ls = min_length_scale, upper_ls = max_length_scale)

prior_rho = prior_rho$par
```

* The minimum length scale is: `r min_length_scale`
* The maximum length scale is: `r max_length_scale`
* The inverse gamma prior parameters for rho (for each of the AccidentYear and DevelopmentLag dimensions are) are: `r round(prior_rho,3)`





# Introduction

## Models Analyzed

This notebook tests key algorithms across the different business lines.  The algorithms tested include

* Multi Company GP on Incremental Loss Ratios with Hurdle Model
* Multi Company GP on Incremental Loss Ratios with Hurdle Model,with individual company lengthscale factor

* Mack Chain Ladder
* Bootstrap Chain Ladder

* Plain GP on Incremental Loss Ratios  
* GP on Incremental Loss Ratios  with hurdle model
* GP on Incremental Loss Ratios  with hurdle model and virtual data points



Since the Chain Ladder does not produce a distribution, the rank and distribution tests cannot be performed.  For all other methods, the distribution tests are performed.

## Stan Programs

This notebook runs Stan programs iteratively.  It is important for speed purposes that each of the Stan programs are pre-compiled. Once the code is compiled, Stan stores a compiled version.  But sometimes this routine fails if the compilation is done as part of a subroutine. Thus it is important that the following code be run once MANUALLY before running the rest of the program.  

```{r}
stanMod <- stan_model(file=paste0(stan_dir, "gp_compound_mean_het_05.stan"))
stanMod <- stan_model(file=paste0(stan_dir,  "gp_hurdle_phiapprox_05.stan"))

stanMod <- stan_model(file=paste0(stan_dir,  "gp_hurdle_multi_co_phiapprox_05.stan"))
stanMod <- stan_model(file=paste0(stan_dir,  "gp_hurdle_multi_co_indiv_rho_05.stan"))

```



# Multi-Company ILR Models

## Multi Company GP on Incremental Loss Ratios with Hurdle Model

  
```{r echo=FALSE, message=FALSE, include=FALSE, results='hide'}

res_gp_inc_hurdle_multi_co = all_business_line_calcs(projection_fun = stan_ilr_multi_co, business_lines = business_lines, 
                                combined_co = TRUE, data = data,
                                iter = iter_multi, warmup=warmup_multi, chains=chains, 
                                stanFile = "gp_hurdle_multi_co_phiapprox_05.stan",
                                prior_eta = 1,  prior_rho = prior_rho, 
                                prior_theta = 1, prior_sigma = 0.5, 
                                prior_alpha = 1)

save(res_gp_inc_hurdle_multi_co , file = paste0(output_dir, "report_algos_gp_inc_hurdle_multi_co_5.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_hurdle_multi_co_5.RData"))

```


```{r}
print_loss_results(res_gp_inc_hurdle_multi_co)
```

## Multi Company GP on Incremental Loss Ratios with Hurdle Model,with individual company lengthscale factor


```{r echo=FALSE, message=FALSE, include=FALSE, results='hide'}

res_gp_inc_hurdle_multi_co_indiv = all_business_line_calcs(projection_fun = stan_ilr_multi_co, business_lines = business_lines, 
                                combined_co = TRUE, data = data,
                                iter = iter_multi, warmup=warmup_multi, chains=chains, 
                                stanFile = "gp_hurdle_multi_co_indiv_rho_05.stan",
                                prior_eta = 1,  prior_rho = prior_rho, 
                                prior_theta = 1, prior_sigma = 0.5, 
                                prior_rho_company = 1)

save(res_gp_inc_hurdle_multi_co_indiv , file = paste0(output_dir, "report_algos_gp_inc_hurdle_multi_co_indiv.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_hurdle_multi_co_indiv.RData"))

```


```{r}
print_loss_results(res_gp_inc_hurdle_multi_co_indiv)
```







# The Single Company Algorithms



## Mack Chain Ladder


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

res_mack_ladder = all_business_line_calcs(business_lines = business_lines, data = data, projection_fun = mack_ladder, 
                                          iter = (iter-warmup)*chains,  include_rank_tests = TRUE)


save(res_mack_ladder , file = paste0(output_dir, "report_algos_mack_ladder_multicompare.RData"))
#load(file = paste0(output_dir, "report_algos_mack_ladder_multicompare.RData"))


```

```{r  echo=FALSE}
print_loss_results(res_mack_ladder)


```


## Bootstrapped Chain Ladder

```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

res_boot_ladder = all_business_line_calcs(business_lines = business_lines, 
                                          data = data, projection_fun = boot_ladder, iter = (iter-warmup)*chains,
                                          include_rank_tests = TRUE)


save(res_boot_ladder , file = paste0(output_dir, "report_algos_boot_ladder_multicompare.RData"))
#load(file = paste0(output_dir, "report_algos_boot_ladder_multicompare.RData"))


```

```{r  echo=FALSE}
print_loss_results(res_boot_ladder)
```



## GP on Incremental Loss Ratios with no constraints on GP using a compound kernel


```{r include=FALSE, echo=FALSE,message=FALSE, results='hide'}

res_gp_inc = all_business_line_calcs(business_lines = business_lines, data = data, projection_fun = stan_incremental_loss_ratio3,  
                                     calc_ranks = T, include_rank_tests = TRUE, verbose = FALSE,
                                     iter = iter, warmup=warmup, chains=chains, adapt_delta = adapt_delta, 
                                     max_treedepth = max_treedepth,
                                     prior_eta = 1, prior_rho = prior_rho, prior_theta = 1, prior_sigma = 0.5, take_logs = FALSE,
                                     mu=0,
                                     formula_x = y_scaled ~ AccidentYear_scaled + DevelopmentLag_scaled -1,
                                     formula_H = y_scaled ~ AccidentYear_scaled + log(DevelopmentLag) -1,
                                     stanFile =  "gp_compound_mean_het_05.stan")


save(res_gp_inc , file = paste0(output_dir, "report_algos_gp_inc_multicompare.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_multicompare.RData"))


```




```{r,echo=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
print_loss_results(res_gp_inc)
```


## GP on Incremental Loss Ratios with Hurdle Model with Vector mu


```{r echo=FALSE, message=FALSE, include=FALSE, results='hide'}

res_gp_inc_hurdle = all_business_line_calcs(business_lines = business_lines, data = data, projection_fun = stan_incremental_loss_ratio3,  
                                     calc_ranks = T, verbose = FALSE,
                                     iter = iter, warmup=warmup, chains=chains, adapt_delta = adapt_delta, 
                                     max_treedepth = max_treedepth,
                                     prior_sigma = .5, mu=10:1 / 10 -.5, prior_rho = prior_rho, 
                                     prior_eta = 1, prior_theta = 1, take_logs = FALSE, 
                                     formula_x = y_scaled ~ AccidentYear_scaled + DevelopmentLag_scaled -1,
                                     formula_H = y_scaled ~ AccidentYear_scaled + log(DevelopmentLag) -1,
                                     stanFile = "gp_hurdle_phiapprox_05.stan")
save(res_gp_inc_hurdle , file = paste0(output_dir, "report_algos_gp_inc_hurdle_multicompare.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_hurdle_multicompare.RData"))

```




```{r,echo=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
print_loss_results(res_gp_inc_hurdle)

```



## GP on Incremental Loss Ratios with Hurdle Model with Vector mu, with Virtual Data 


```{r echo=FALSE, message=FALSE, include=FALSE, results='hide'}
res_gp_inc_hurdle_virtual = all_business_line_calcs(business_lines = business_lines, data = data, 
                                                    projection_fun = stan_incremental_loss_ratio3,  
                                     calc_ranks = T, verbose = FALSE,
                                     iter = iter, warmup=warmup, chains=chains, adapt_delta = adapt_delta, 
                                     max_treedepth = max_treedepth,
                                     #prior_rho = c(5.566019, 5.570317), 
                                     prior_rho = prior_rho, 
                                     #prior_sigma = 1, #
                                     prior_sigma = .5, 
                                     mu=10:0 / 10 -.5, 
                                     prior_eta = 1, prior_theta = 1, take_logs = FALSE, 
                                     formula_x = y_scaled ~ AccidentYear_scaled + DevelopmentLag_scaled -1,
                                     formula_H = y_scaled ~ AccidentYear_scaled + log(DevelopmentLag) -1,
                                      virtual_data = T, virtual_point = -.3, virtual_accident_year = 1988:1997, virtual_development_lag = 11,
                                     stanFile = "gp_hurdle_phiapprox_05.stan")
save(res_gp_inc_hurdle_virtual , file = paste0(output_dir, "report_algos_gp_inc_hurdle_virtual.RData"))
#load( file = paste0(output_dir, "report_algos_gp_inc_hurdle_virtual.RData"))

```




```{r,echo=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
print_loss_results(res_gp_inc_hurdle_virtual)

```


